Luminance constancy under fixed geometry

### 2018-02-08 Draft2/draft2.tex
more edits

The computational work reviewed above is based on formulating an explicit model of the image formation process and the statistical properties of natural images, and leveraging these formulations to obtain an appropriate ``inverse optics" algorithm.  An alternative approach, which has been less-well explored, is to use supervised machine learning techniques to develop mappings between input images and stable color descriptors.  Along these lines, Barron \cite{barron2015convolutional} recently developed a color balancing algorithm based on training a neural network to detect a uniform illuminant in an image by matching image templates in the chromaticity space. The use of supervised learning to understand human perceptual capabilities has enjoyed success in domains outside of color vision. Geisler and colleagues, for example, have developed a technique they refer to as accuracy maximization analysis (AMA, \citeNP{geisler2009optimal}), which learns linear filters whose output carries information that is optimized for particular perceptual tasks. This technique has provided computations that efficiently estimate speed \cite{burge2015optimal}, focus \cite{burge2011optimal} and disparity \cite{burge2014optimal} 
% JOHANNES ARE THERE ANY OTHER PAPERS YOU WOULD LIKE TO CITE
from image patches taken from natural scenes, and that in addition provide excellent models of human's ability to make corresponding estimates from the same stimuli.  In this paper, we apply AMA, as well as comparison supervised learning algorithms to a special case of the color constancy problem (luminance constancy) and evaluate how well it performs for naturalistic image input.

@@

In this work, we are interested in how well luminance constancy may be achieved using reasonable information that is local to the image region containing the target patch. 
% MAYBE SAY A LITTLE MORE ABOUT WHY - THIS IS A PRINCIPLED STANCE TAKEN IN MOST OF JOHANNES WORK.  OR MAYBE DEFER THAT TO THE DISCUSSION.
To this end, we crop a 51x51 pixel portion of the full rendered multispectral image that contains the target object and use this for further analysis (\ref{fig:croppedImage}).
Fig. \ref{fig:studiedCases} shows examples of such cropped image at 3 luminance levels for the three cases considered in this paper.
Next, the cropped multispectral images are used to simulate retinal cone responses using \href{https://github.com/isetbio}{Isetbio} software. The key steps we model are the size of the pupil, optical blurring (including axial chromatic aberration) in the formation of the retinal image, spatial sampling by an interleaved mosaic of long (L), middle (M)  and short (S) wavelength-sensitive cones, and the Poisson nature of photopigment isomerization.
We use the multispectral cropped image (Fig.~\ref{fig:croppedImage}) as the optical stimulus to simulate the response of the cones in the retina. The cropped image patch is assumed to cover a $1^{\circ}$ field of view at a distance of 1 m. The retinal cone mosaic is modeled as a square grid with $51 \times 51$ cones in the L:M:S ratio 0.6:0.3:0.1. The response of the cones are taken as the number of isomerizations in the cone segment ({\it over what time ??}).

@@

The isomerizations of the cones in the retinal mosaic represent the first steps of visual processing.  To capture key properties of post-receptoral processing, contrast coding (REFS) and contrast normalization (REFS),
% Ask Johannes what he'd use as references for these broad principles
we processed the isomerizations as follows:
First, we created a representation for each cone class on the full $51 \times 51$ pixel grid, we interpolated the responses of the cones in each class across the grid, using bilinear interpolation separately for each cone class.
% DOUBLE CHECK THAT THIS WAS DONE WITH SOMETHING LIKE INTERP2 WITH THE LINEAR OPTION.
This gives us a $51 \times 51$ cone response image for each cone class.
The three $51 \times 51$ sets of cone responses corresponding to the L, M and S cones were then reshaped into a single $7803 \times 1$ vector ($7803 = 51 \times 51 \times 3$). For case 1 and 3 we used this as the input for the supervised methods. For case 3 and 4, to model the contrast coding and contrast adaptation, we convert this vector to a contrast representation by subtracting out the mean and then dividing by it (contrast coding), followed by normalizing the vector to have unit length (contrast normalization).
The cone-response vectors corresponding to the N images for each case are concatenated column-wise to form a contrast normalized matrix $C_{7803\times N}$. The corresponding labels are represented by the row vector $Y_{1\times N}$.

@@

\subsubsection*{Accuracy maximization analysis}
% MAYBE ADD SOME INTRO TEXT THAT SAYS THAT FIRST ONE FINDS THE FILTERS AND THEN PROJECTS ONTO THEM, THEN DOES BAYESIAN DECODING OF THE RESPONSES.  BREAK INTO TWO SUBSECTIONS FOR CLARITY.

To construct an ideal observer model for our task, two steps are required. First, we use AMA to learn the filters that are optimal for the task. Second, we determine how to optimally decode the task relevant variable from the filter responses to arbitrary stimuli.

{\bf Learning optimal filters:} AMA \cite{geisler2009optimal,burge2017accuracy} is a task specific Bayesian method for dimensionality reduction. When provided with a labeled training set, a response noise model, and a cost function, AMA returns the linear receptive fields in rank order that are optimal for performing a task. The optimal filters that are returned by AMA are the filters whose noisy responses to the stimuli, as given by the specified response model, minimizes the cost function over the training set when optimally decoded.

Let us assume that there are $N_{\nu}$ possible values of the target object luminance levels ($\{Y_k: k\in[1,N_{\nu}] \}$). We represent the $l$'th image corresponding to the $k$'th level as $s_{kl}$. Let us assume that there is a set of linear filters {\textbf{\textit f}} which when acting on $s_{kl}$ gives the response ${\textbf{\textit R}}_{kl}$. Let's say we have a decoder $g$, that uses the information of the response noise model, to return the Bayes optimal estimate of the luminance level ($\hat{Y}_k$) corresponding to the stimuli $s_{kl}$. Given this estimate, one can define a cost function $\mathcal{C}_{kl}(Y_{kl},\hat{Y}_{kl})$ corresponding to specific response ${\textbf{\textit R}}_{kl}$ and an avearge cost $\bar{\mathcal{C}}_{kl}(Y_{kl},\hat{Y}_{kl})$ corresponding to the average response $r_{kl} = \langle{\textbf{\textit R}}_{kl}\rangle$.

Accuracy maximization analysis (AMA) finds the set of optimal filters ${\textbf{\textit f}}^{\rm opt}$ that minimize the average cost function over the entire dataset $\{s_{kl}: k\in[1,N_{\nu}], l\in[1,N_k]\}$. Here $N_{k}$ is the number of stimuli corresponding to the label $Y_k$. Thus,
\begin{align}
{\textbf {\textit f} }^{\rm opt} = \argmin_{\textbf {\textit f} } \sum_{kl} \bar{\mathcal{C}}_{kl}.
\label{eq:fopt}
\end{align}

AMA employs the Bayes optimal decoder ($g$) to estimate the optimal estimate. The Bayesian decoder computes the posterior probability of the label $p(Y|{\textbf {\textit R}})$ and reads out the optimal estimate $\hat{Y}^{\rm opt}$ from the posterior as the value that minimizes the cost. By Bayes' theorem, the posterior probability can be written as:
\begin{align}
p(Y_k|{\textbf {\textit R} }_{kl}) = \frac{p({\textbf {\textit R} }_{kl}|Y_k)p(Y_k)}{\sum\limits_{i=1}^{N_{\nu}}{p({\textbf {\textit R} }_{kl}|Y_i)p(Y_i)}}.
\label{eq:posterior}
\end{align}
Since, the conditional probability of the filter response is:
\begin{align}
p({\textbf {\textit R} }_{kl}|Y_k) = \sum\limits_{m=1}^{N_k}p({\textbf {\textit R} }_{kl}|{\textbf {\textit s} }_{km})p({\textbf {\textit s} }_{km}|Y_k),
\label{eq:condProb}
\end{align}
using Eq.~\ref{eq:condProb} in Eq.~\ref{eq:posterior} we get:
\begin{align}
p(Y_k|{\textbf {\textit R} }_{kl}) = \frac{\left[\sum\limits_{m=1}^{N_k}p({\textbf {\textit R} }_{kl}|{\textbf {\textit s} }_{km})p({\textbf {\textit s} }_{km}|Y_k)\right]p(Y_k)}{\sum\limits_{i=1}^{N_{\nu}}{\left[\sum\limits_{j=1}^{N_j}p({\textbf {\textit R} }_{il}|{\textbf {\textit s} }_{ij})p({\textbf {\textit s} }_{kj}|Y_i)\right]p(Y_i)}}
\label{eq:finalPosterior}
\end{align}
Since, $p({\textbf {\textit s} }_{km}|Y_k) = \frac{1}{N_k}$, and $p(Y_k)=\frac{N_k}{N}$, where $N = \sum\limits_{k=1}^{N_{\nu}}N_{k}$ is total number of samples in the training set, Eq.~\ref{eq:finalPosterior} simplifies to
\begin{align}
p(Y_k|{\textbf {\textit R} }_{kl}) = \frac{\left[\sum\limits_{m=1}^{N_k}p({\textbf {\textit R} }_{kl}|{\textbf {\textit s} }_{km}) \frac{1}{N_k}\right]\frac{N_k}{N}}{\sum\limits_{i=1}^{N_{\nu}}{\left[\sum\limits_{j=1}^{N_j}p({\textbf {\textit R} }_{il}|{\textbf {\textit s} }_{ij}) \frac{1}{N_i}\right]\frac{N_i}{N}}} = \frac{\sum\limits_{m=1}^{N_k}p({\textbf {\textit R} }_{kl}|{\textbf {\textit s} }_{km})}{\sum\limits_{i=1}^{N_{\nu}}{\sum\limits_{j=1}^{N_j}p({\textbf {\textit R} }_{il}|{\textbf {\textit s} }_{ij})}}.
\label{eq:simplifiedPosterior}
\end{align}
Now if the cost associated with the latent variable $Y_k$ and the estimate $\hat{Y}_k$ is $\gamma (Y_k,\hat{Y}_k)$, we have  
\begin{align}
\mathcal{C}_{kl} = \sum\limits_{u=1}^{N_{\nu}}\gamma(Y_u,\hat{Y}_u)p(X_u|{\textbf{\textit R}}_{kl}).
\end{align}
The mean cost associated with the training set is $\bar{\mathcal{C}} = \frac{1}{N} \sum\limits_{k,l}\bar{\mathcal{C}}_{kl} = \frac{1}{N} \sum\limits_{k,l}\langle \mathcal{C}_{kl}\rangle_{{\textbf{\textit R}}_{kl}}$. The optimal AMA filters are obtained by minimizing $\bar{\mathcal{C}}$ over the filter set ${\textbf {\textit f} }$ (Eq.\ref{eq:fopt}). The filters are obtained using gradient descent.

In this work, we assume the noisy response model to be a multivariate gaussian. For a stimuli $s_{kl}$, the noisy response vector ${\textbf{\textit R}}_{kl} = \left[{\textbf{\textit R}}_{kl,1},{\textbf{\textit R}}_{kl,2},\dots,{\textbf{\textit R}}_{kl,q}\right]$ to the set of filters ${\textbf{\textit f}} = \left[{\textbf{\textit f}}_{1},{\textbf{\textit f}}_{2},\dots,{\textbf{\textit f}}_{q}\right]$ is given by $\mathcal{N}({\textbf{\textit r}}_{kl}, \Sigma)$, with mean response vector ${\textbf{\textit r}}_{kl} = \left[{\textbf{\textit r}}_{kl,1},{\textbf{\textit r}}_{kl,2},\dots,{\textbf{\textit r}}_{kl,q}\right] = \left[{\textbf{\textit f}}_{1}\cdot s_{kl},{\textbf{\textit f}}_{2}\cdot s_{kl}\dots,{\textbf{\textit f}}_{q}\cdot s_{kl}\right]$ and the covariance matrix as the diagonal matrix $\Sigma$ with variances $\left(\sigma_1^2, \sigma_2^2,\dots, \sigma_q^2 \right)$. We assume $\sigma_i^2 = \alpha |r_i| + \sigma_0^2$, $|{\textbf{\textit f}}_{kl,i}|=1$ and $s_{kl}$ to be contrast normalized, i.e., ${|s_{kl}|=1}$.

>>>>

The computational work reviewed above is based on formulating an explicit model of the image formation process and the statistical properties of natural images, and leveraging these formulations to obtain an appropriate ``inverse optics" algorithm.  An alternative approach, which has been less-well explored, is to use supervised machine learning techniques to develop mappings between input images and stable color descriptors.  Along these lines, Barron \cite{barron2015convolutional} recently developed a color balancing algorithm based on training a neural network to detect a uniform illuminant in an image by matching image templates in the chromaticity space. The use of supervised learning to understand human perceptual capabilities has enjoyed success in domains outside of color vision. Geisler and colleagues, for example, have developed a technique they refer to as accuracy maximization analysis (AMA, \citeNP{geisler2009optimal}), which learns linear filters whose output carries information that is optimized for particular perceptual tasks. This technique has provided computations that efficiently estimate speed \cite{burge2015optimal}, focus error \cite{burge2011optimal}, and disparity \cite{burge2014optimal} 
from image patches taken from natural scenes, and that in addition provide excellent models of human's ability to make corresponding estimates from the same stimuli.  In this paper, we apply AMA, as well as comparison supervised learning algorithms to a special case of the color constancy problem (luminance constancy) and evaluate how well it performs for naturalistic image input.

@@

In this work, we are interested in how well luminance constancy may be achieved using information that is local to the image region containing the target object. 
To analyze the local image information, we cropped and processed the image in the immediate neighborhood of the target object such that each patch represented a one degree of visual angle, approximately the size of a receptive field in early visual cortex (\ref{fig:croppedImage}).
Fig. \ref{fig:studiedCases} shows examples of such cropped image at 3 luminance levels for the three cases considered in this paper.
Next, the cropped multispectral images are used to simulate retinal cone responses using \href{https://github.com/isetbio}{Isetbio} software. The key steps we model are the size of the pupil, optical blurring (including axial chromatic aberration) in the formation of the retinal image, spatial sampling by an interleaved mosaic of long (L), middle (M)  and short (S) wavelength-sensitive cones, and the Poisson nature of photopigment isomerization.
We use the multispectral cropped image (Fig.~\ref{fig:croppedImage}) as the optical stimulus to simulate the response of the cones in the retina. The cropped image patch is assumed to cover a $1^{\circ}$ field of view at a distance of 1 m. The retinal cone mosaic is modeled as a square grid with $51 \times 51$ cones in the L:M:S ratio 0.6:0.3:0.1. The response of the cones are taken as the number of isomerizations in the cone segment ({\it over what time ??}).

@@

The cone isomerizations represent the first step of visual processing.  To capture key properties of post-receptoral processing, such as contrast normalization (REFS),
% Heeger 1992, Albrecht and Geisler 1991 (contrast normalization)
% Carindini Heeger 2012 (Review on contrast norm)
we processed the isomerizations as follows:
First, we created a representation for each cone class on the full $51 \times 51$ pixel grid, we interpolated the responses of the cones in each class across the grid, using bilinear interpolation separately for each cone class.
% DOUBLE CHECK THAT THIS WAS DONE WITH SOMETHING LIKE INTERP2 WITH THE LINEAR OPTION.
This gives us a $51 \times 51$ cone response image for each cone class.
The three $51 \times 51$ sets of cone responses corresponding to the L, M and S cones were then reshaped into a single $7803 \times 1$ vector ($7803 = 51 \times 51 \times 3$). For case 1 and 3 we used this as the input for the supervised methods. For case 3 and 4, to model the contrast coding and contrast adaptation, we convert this vector to a contrast representation by subtracting out the mean and then dividing by it (contrast coding), followed by normalizing the vector to have unit length (contrast normalization).
The cone-response vectors corresponding to the N images for each case are concatenated column-wise to form a contrast normalized matrix $C_{7803\times N}$. The corresponding labels are represented by the row vector $Y_{1\times N}$.

@@

\subsubsection*{Accuracy maximization analysis}
We perform two steps to decode the luminance of the target object from the cone isomerization. First, we use AMA to learn the linear filters that are optimal for estimating the target object luminance. Second, we determine how to optimally decode the luminance from the response of these filters to arbitrary stimuli.

{\bf Learning optimal filters using AMA:} AMA \cite{geisler2009optimal,burge2017accuracy}
% add Jaini burge 2017
is a task specific Bayesian method for dimensionality reduction. When provided with a labeled training set, a receptive field response model, a decoder that uses these responses to estimate the stimulus label, and a cost function, AMA returns the linear receptive fields (in rank order) that are optimal for performing a task. The optimal receptive fields that are returned by AMA are those whose responses to the stimuli minimizes the cost function over the training set when optimally decoded.

More concretely, let us assume that there are $N_{\nu}$ possible values of the target object luminance levels ($\{Y_k: k\in[1,N_{\nu}] \}$) and the $l$'th image corresponding to the $k$'th luminance level is represented as $s_{kl}$. Also, assume that there is a set of linear receptive fields {\textbf{\textit f}} which when acting on $s_{kl}$ gives the response ${\textbf{\textit R}}_{kl}$. Also assume there is a decoder $g$, with the knowledge of the response model and of every stimulus in the training set, that returns the Bayes' optimal estimate of the luminance level ($\hat{Y}_k$) of the target object in each stimulus $s_{kl}$. The crux computation for obtaining this estimate is the computation of the posterior probability of the luminance given an observed response, $P(Y|\textbf{R}_{kl})$ \cite{geisler2009optimal,burge2017accuracy}
% add Jaini burge 2017
. With this luminance estimate, we calculate the cost given the cost function $\mathcal{C}_{kl} = \gamma(Y_{kl},\hat{Y}_{kl})$. Finally, AMA searches for the set of optimal receptive fields ${\textbf{\textit f}}^{\rm opt}$ that minimize the cost over the entire training set $\{s_{kl}: k\in[1,N_{\nu}], l\in[1,N_k]\}$ where $N_{k}$ is the number of stimuli corresponding to the label $Y_k$. Thus,
\begin{align}
{\textbf {\textit f} }^{\rm opt} = \argmin_{\textbf {\textit f} } \sum_{k=1}^{N_{\nu}}  \sum_{l=1}^{N_k} {\mathcal{C}}_{kl}.
\label{eq:fopt}
\end{align}

{\bf Estimating luminance:} 
Once the optimal receptive fields have been learned, a general decoder must be used to estimate the luminance. Recall that the decoder used to learn the receptive fields required the knowledge of every stimulus in the training set. To obtain the general decoder, first we use the AMA optimal receptive fields and the training dataset to find the conditional distributions of the AMA receptive field responses $P({\textbf {\textit R} }|Y_i)$. Then, we approximate these with multivariate Gaussian distributions $P({\textbf {\textit R} }|Y_i) = \mathcal{N}\left(\mu(Y_i),\Sigma(Y_i) \right)$. To estimate the luminance of the target object in the test stimulus, we use Bayes' rule to obtain the posterior distribution $P(Y|{\textbf {\textit R} })$. The optimal estimate is the luminance value that minimizes the cost function.
